---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("utils/utils.R")
```

## Overview

This repository accompanies the manuscript _"Comprehensive mass spectrometry workflows to systematically elucidate
transformation processes of organic micropollutants: a case study on photodegradation of four pharmaceuticals"_ and
contains the [R] and [patRoon] based data processing scripts and algorithms that were used to elucidate transformation
products from photodegradation experiments.

The repository layout is as follows:

* `workflow.R`: the main processing script
* `workflow/` directory: script files that implement the main workflow
* `utils/` directory: scripts with utility functions used throughout the workflow
* `report_pd` directory: all support files needed for automatic report generation
* `misc/` directory: miscellaneous workflow input files (suspect lists, sample info etc)
* `output/` directory: output files generated throughout the workflow
    * specifically: `suspects_TPs-compounds.csv` and `suspects_TPs-formulas.csv`: suspect lists generated during TP suspect screening workflows
* `renv.lock`: an [renv] lockfile to recreate a reproducible R environment

## Used software

### General

Software       | Version | URL
-------------- | ------- | -------------------------------------
R              | 4.2.2   | https://www.r-project.org/
DataAnalysis   | 4.4     | https://www.bruker.com/
OpenMS         | 2.7     | https://openms.de/
GenForm        | embededded in patRoon | https://sourceforge.net/projects/genform/
MetFrag        | 2.5     | http://ipb-halle.github.io/MetFrag/
BioTransformer | 3.0.0   | http://biotransformer.ca/
Chemical Transformation Simulator | Accessed online July 19th 2024 | https://qed.epa.gov/cts/
MolView        | 2.4     | https://molview.org/
CDK Depict     | --      | https://www.simolecule.com/cdkdepict/depict.html
OpenBabel      | 3.1.1   | https://github.com/openbabel/openbabel
Inkscape       | 1.2.2   | https://inkscape.org/

### Used R packages

```{r pkgTab,message=FALSE}
knitr::kable(getAllRPackages())
```

[R]: https://www.r-project.org/
[patRoon]: https://rickhelmus.github.io/patRoon/
[renv]: https://rstudio.github.io/renv
